---
title: "Seminar 3: Sandbox"
author: "Max Glonek"
date: "Trimester `r if(as.integer(format(Sys.time(),'%m'))<5){1}else if(as.integer(format(Sys.time(),'%m'))<9){2}else{3}`, `r format(Sys.time(),'%Y')`"
output:
  bookdown::pdf_document2:
    toc: no
    includes:
      in_header: header.tex
  bookdown::html_document2: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
 echo = TRUE, 
 fig.width = 6, 
 fig.asp = 0.618, 
 out.width = "70%",
 fig.align = "center", 
  root.dir = '../'
)
pacman::p_load(tidyverse,tidymodels,bookdown)
```

# Try it Yourself: Add a Figure

```{r longleyPlot, fig.cap='Scatterplot of Employed vs GNP for the longley dataset', echo=FALSE}
data("longley")
plot(Employed ~ GNP, data = longley, col='#3fbf7f', pch=20, cex=2)
abline(lm(Employed ~ GNP, data = longley), col = '#bf7fff', lwd = 2)
```

Figure \@ref(fig:longleyPlot) shows a strong, positive, linear relationship between number of employed people and Gross National Product.

# Try it Yourself: Add a Table

```{r toothTab, echo=FALSE}
data("ToothGrowth")
ToothGrowth %>%
  count(supp) %>%
  knitr::kable(caption = 'TootGrowth factors.')
```

Table \@ref(tab:toothTab) shows a balanced study design in the ToothGrowth dataset.

# Try it Yourself: Formatting Mathematics

# Try it Yourself: Formatting Environments

# Try it Yourself: Displaying Code You Don't Want to Run