---
title: "STATS 7022 - Data Science PG Assignment 2"
author: "Dang Thinh Nguyen"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE, 
                      warning = FALSE,
                      fig.dim=c(6,5),
                      fig.align='center')
```

# **Question 1: Data Analysis**

## 1 Load libraries

```{r}
pacman::p_load(tidyverse, bookdown, readr, dplyr, tidymodels, pROC, knitr, kableExtra)
```

## 2 Load data

```{r}
# Read in the data
data <- readRDS('./diamonds.rds')

# Display the first 10 lines of the data
data %>% head(10)
```

## 3 Data Cleaning and Pre-processing

### 3.1 Remove Missing Prices and Volumes

```{r}
# Check missing values
sum(is.na(data))

# Remove missing values
data_cleaned <- na.omit(data)

# Check missing values after removing
sum(is.na(data_cleaned))
```

### 3.2 Derive Cut, Colour, and Clarity

### 3.2.1 Cut

```{r}
# Create 'cut' column based on the first character of 'c.grade'
data_cleaned <- data_cleaned %>%
  mutate(
    cut = ifelse(substr(c.grade, 1, 1) == '1', 'premium', 'ideal')
  )

# Create table of 'cut'
cut_table <- data_cleaned %>% count(cut)

# Display the table
kable(cut_table, col.names = c('cut', 'count'),
      caption = 'Table 3.1: Number of diamonds, by cut quality.') %>%
  kable_styling(full_width = T)
```

### 3.2.2 Colour

```{r}
# Create 'colour' column based on the second character of 'c.grade'
data_cleaned <- data_cleaned %>%
  mutate(
    colour = substr(c.grade, 2, 2)
  )

# Create table of 'cut'
colour_table <- data_cleaned %>% count(colour)

# Display the table
kable(colour_table, col.names = c('colour', 'count'),
      caption = 'Table 3.2: Number of diamonds, by colour.') %>%
  kable_styling(full_width = T)
```

### 3.2.3 Clarity

```{r}
# Create a map for clarity
clarity_map <- c('0' = 'IF',
                 '1' = 'VVS1',
                 '2' = 'VVS2',
                 '3' = 'VS1',
                 '4' = 'VS2',
                 '5' = 'SI1',
                 '6' = 'SI2',
                 '7' = 'I1')

# Create 'clarity' column based on the third character of 'c.grade'
data_cleaned <- data_cleaned %>%
  mutate(clarity = clarity_map[substr(c.grade, 3, 3)])

# Create table of 'cut'
clarity_table <- data_cleaned %>% count(clarity)

# Display the table
kable(clarity_table, col.names = c('clarity', 'count'),
      caption = 'Table 3.3: Number of diamonds, by clarity.') %>%
  kable_styling(full_width = T)
```

### 3.3 Select variables

```{r}
# Select variables
data_2 <- data_cleaned %>%
  dplyr::select(cut, price, volume)

# Display the first 10 lines of the data
data_2 %>% head(10)
```

### 3.4 Convert Cut to categorical

```{r}
# Convert 'cut' to factor
data_2 <- data_2 %>% 
  mutate(cut = as.factor(cut))

# Display the first 10 lines of the data
data_2 %>% head(10)
```

## 4 Model

### 4.1 Logistic regression model with cut as the response variable and price as the predictor

```{r}
# Preprocessor 
recipe_cp <- recipe(cut ~ price, data = data_2) %>%
  step_normalize()

# Logistic regression model
log_reg_model <- logistic_reg() %>%
  set_mode('classification') %>%
  set_engine('glm')

# Workflow
wf_cp <- workflow() %>%
  add_recipe(recipe_cp) %>%
  add_model(log_reg_model)
wf_cp
```

### 4.2 Logistic regression model with cut as the response variable and volume as the predictor

```{r}
# Preprocessor 
recipe_cv <- recipe(cut ~ volume, data = data_2) %>%
  step_normalize()

# Workflow
wf_cv <- workflow() %>%
  add_recipe(recipe_cv) %>%
  add_model(log_reg_model)
wf_cv
```

## 5 ROC Curves

```{r data_2, fig.cap='Figure 5.1: ROC Curves for logistic regression models.'}
# Fit the models
fit_cp <- fit(wf_cp, data = data_2)
fit_cv <- fit(wf_cv, data = data_2)

# Predict probabilities and add true class to predictions
pred_price <- predict(fit_cp, data_2, type = "prob") %>% 
  bind_cols(data_2 %>% dplyr::select(cut))
pred_volume <- predict(fit_cv, data_2, type = "prob") %>% 
  bind_cols(data_2 %>% dplyr::select(cut))

# Calculate ROC curve metrics
roc_price <- pred_price %>% 
  roc_curve(truth = cut, .pred_ideal)
roc_volume <- pred_volume %>% 
  roc_curve(truth = cut, .pred_ideal)

# Add model names to each table
roc_price <- roc_price %>% 
  mutate(model = "Price")
roc_volume <- roc_volume %>% 
  mutate(model = "Volume")

# Combine the two tables
roc_data <- bind_rows(roc_price, roc_volume)

# Plot ROC curves
ggplot(roc_data, aes(x = 1 - specificity, y = sensitivity, color = model)) +
  geom_line() +
  geom_abline(linetype = "dashed") +
  labs(x = "1 - specificity",
       y = "sensitivity") +
  theme_minimal() +
  theme(panel.border = element_rect(color = "black", fill = NA, size = 0.5))
```